title: "model_transfusion"
author: "Lisa Eisler"
date: "9/30/2020"
output: github_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(viridis)
```

## Data import 

```{r, message=FALSE, warning=FALSE}
nsqipspine_1618 = read_csv("./data/nsqipspine1618.csv")
```

```{r, message=FALSE, warning=FALSE}
nsqipspine_tidy = nsqipspine_1618 %>%
  mutate_if(is.numeric, ~replace(., . == -99, NA)) 
```

##Sex
```{r, message=FALSE, warning=FALSE}
xtabs(~ sex + perioptransfusion, data = nsqipspine_tidy)
```

no cells with counts<=5
```{r, message=FALSE, warning=FALSE}
tbl<-xtabs(~ sex + perioptransfusion, data = nsqipspine_tidy)
chisq.test(tbl) 
```
Chi-square test reveals a significant negative association between female sex and perioperative transfusion (p<.05).

##cardiac
```{r, message=FALSE, warning=FALSE}
xtabs(~ cardiac + perioptransfusion, data = nsqipspine_tidy)
```

no cells with counts<=5
```{r, message=FALSE, warning=FALSE}
tbl<-xtabs(~ cardiac + perioptransfusion, data = nsqipspine_tidy)
chisq.test(tbl) 
```
Chi-square test reveals a significant positive association between cardiac risk factors and perioperative transfusion (p<.05).
