title: "model_transfusion"
author: "Lisa Eisler"
date: "9/30/2020"
output: github_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)
library(viridis)
```


## Data import 

```{r, message=FALSE, warning=FALSE}
nsqipspine_1618 = read_csv("./data/nsqipspine_1618.csv")
```

## Initial Data Cleaning & Selection of a subset of data to examine the purpose of the study 

```{r, message=FALSE, warning=FALSE}
spine_tidy = nsqipspine_1618 %>%
  mutate_if(is.numeric, ~replace(., . == -99, NA)) %>%
  mutate(
    age_years = age_days/365.25,
    height = height*2.54,
    weight = weight/2.2) %>%
  mutate(
    bmi = weight/((height/100)*(height/100)),
    asa_status = case_when(
      asaclas == "ASA 1 - No Disturb" ~ "1",
      asaclas == "ASA 2 - Mild Disturb" ~ "2",
      asaclas == "ASA 3 - Severe Disturb" ~ "3",
      asaclas == "ASA 4 - Life Threat" ~ "4",
      asaclas == "None assigned" ~ "NA"),
    home_discharge = case_when(
      dischdest == "Expired" ~ "FALSE",
      dischdest == "Facility Which was Home" ~ "TRUE",
      dischdest == "Home" ~ "TRUE",
      dischdest == "Rehab" ~ "FALSE",
      dischdest == "Separate Acute Care" ~ "FALSE",
      dischdest == "Skilled Care, Not Home" ~ "FALSE",
      dischdest == "Unknown" ~ "NA",
      dischdest == "Unskilled Facility Not Home" ~ "FALSE",
      dischdest == "NULL" ~ "NA"),
    level_13 = case_when(
      prncptx == "ARTHRODESIS, ANTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 2 TO 3 VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, ANTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 4 TO 7 VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, ANTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 8 OR MORE VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, POSTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; UP TO 6 VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, POSTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 7 TO 12 VERTEBRAL SEGMENTS" ~ "FALSE",
      prncptx == "ARTHRODESIS, POSTERIOR, FOR SPINAL DEFORMITY, WITH OR WITHOUT CAST; 13 OR MORE VERTEBRAL SEGMENTS" ~ "TRUE"))  %>% 
  filter(home_discharge != "NA") %>% 
  select(pufyear_x:ped_spn_post_neurodeftype, age_years, sex, height, weight, bmi, ethnicity_hispanic, race, asa_status, transt, ventilat, asthma, hxcld, oxygen_sup, crf, impcogstat, seizure, nutr_support, hemodisorder, level_13, optime, tothlos, d_opto_dis, death30yn, supinfec, wndinfd, orgspcssi, dehis, oupneumo, pulembol, renainsf, urninfec, cszre, neurodef, cdarrest, othbleed, bleed_ml_tot, othcdiff, othsysep, unplannedreadmission1, reoperation, dischdest, home_discharge)
```

## Analysis and models
Our analysis goal is to develop a model of possible associations between pre-surgical indicators and the outcome of transfusion. We use crosstabs and chisquare tests to examine associations between each of the indicators and the outcome (Preliminary analysis) and logistic regression to build our models.

Data processing, elimination of missing data, and data selection for the analysis. 
```{r, message=FALSE, warning=FALSE}
model_data = spine_tidy %>% 
  mutate(
    sex1=case_when(
    sex=="Male"~"0",
    sex=="Female"~"1"),
    sex1=as.factor(sex1),
    race = case_when(
    race == "American Indian,Alaskan Native" ~ "1",
    race == "Asian" ~ "2",
    race == "Black or African American" ~ "3",
    race == "Native Hawaiian or Other Pacific Islander" ~ "4",
    race == "White" ~ "5",
    race == "Unknown Not Reported" ~ "6"),
    race1=as.factor(race),
    asa_level = ifelse(asa_status >2, 1, 0),
    asa_level=as.factor(asa_level),
    ventilat1=case_when(
    ventilat=="No"~"0",
    ventilat=="Yes"~"1"),
    ventilat1=as.factor(ventilat1), 
  asthma1=case_when(
    asthma=="No"~"0",
    asthma=="Yes"~"1"),
  asthma1=as.factor(asthma1),
  hxcld1=case_when(
    hxcld=="No"~"0",
    hxcld=="Yes"~"1"),
  hxcld1=as.factor(hxcld1),
  oxygen_sup1=case_when(
    oxygen_sup=="No"~"0",
    oxygen_sup=="Yes"~"1"),
  oxygen_sup1=as.factor(oxygen_sup1),
  seizure1=case_when(
    seizure=="No"~"0",
    seizure=="Yes"~"1"),
  seizure1=as.factor(seizure1),
  nutr_support1=case_when(
    nutr_support=="No"~"0",
    nutr_support=="Yes"~"1"),
  nutr_support1=as.factor(nutr_support1),
hemodisorder1=case_when(
    hemodisorder=="No"~"0",
    hemodisorder=="Yes"~"1"),
  hemodisorder1=as.factor(hemodisorder1),
impcogstat1=case_when(
    impcogstat=="No"~"0",
    impcogstat=="Yes"~"1"),
  impcogstat1=as.factor(impcogstat),
 level_13=as.factor(level_13),
 home_discharge1=case_when(
   home_discharge=="TRUE"~"1",
   home_discharge=="FALSE"~"0"),
home_discharge1=as.factor(home_discharge1), 
  crf1 = case_when(
      crf == "Major cardiac risk factors" ~ "TRUE",
      crf == "Severe cardiac risk factors" ~"TRUE",
      crf == "Minor cardiac risk factors" ~ "FALSE",
      crf == "No cardiac risk factors" ~ "FALSE"), 
  crf1=as.factor(crf1),
  transt1 = case_when(
      transt == "Chronic care Rehab/Intermediate Care/Spinal Cord" ~ "FALSE",
      transt == "Transferred from outside hospital (NICU, PICU, Inpatient on General floor, Adult" ~"FALSE",
      transt == "other" ~ "FALSE",
      transt == "Admitted from home/clinic/doctor's office" ~ "TRUE",
      transt == "Admitted through ER, including outside ER with direct hospital admission" ~ "TRUE"),
  transt1=as.factor(transt1))%>%
  select(age_years, bmi, weight, height, sex1, transt1, crf1, race1, asa_level, ventilat1, asthma1,impcogstat1, hxcld1, oxygen_sup1, seizure1, nutr_support1, hemodisorder1, level_13, home_discharge1) 
```
Checking the data with quick descriptives
```{r, message=FALSE, warning=FALSE}
summary(model_data)
```

##Preliminary Analysis Crosstabs & Chisqaure tests of association of each of the indicators with the outcome.
Two-way contingency tables of categorical outcome and predictors identified. Ensure that there no cells with counts <=5 and compute Chi-square tests to examine the association of each of the indicators to the outcome. 

##Sex
```{r, message=FALSE, warning=FALSE}
xtabs(~ home_discharge1+ sex1, data = model_data)
```
no cells with counts<=5



Selection of variables for the Full Model
Initially there were XX pre-operative indicators to be considered for the main outcome `transfusion` To select the variables for our full model we used the following criteria:

1) From groups of variables that were highly correlated with each other, we selected one variable that was supported from the literature. For example, we used weight instead of height and bmi, as height is a non reliable measure for individuals with CP and has a lot of missing values in the dataset. 

2) From the preliminary analysis (crosstabs) we identified and eliminated variables that presented with cell counts <=5. 

3) We finally eliminated variables such as `ethnicity` because there was a large number of patients who self-identified as "Other" and it is unclear how to use "Other" to create preditions. 

Therefore, our full regression model consisted of xx variables. 

##Logistic Regression Models
Full Model
```{r, message=FALSE, warning=FALSE}
mylogit <- glm(home_discharge1 ~ age_years+ weight+ sex1 + crf1+asthma1 +impcogstat1+ hxcld1+ seizure1 + nutr_support1 + hemodisorder1+ level_13, data = model_data, family = "binomial")
summary(mylogit)
```